---
title: "ERRNI Notebook"
output: html_notebook
---

Notebook version of original analysis scripts for Frizelle et al.
The original scripts were uploaded to OSF around Dec 2016 https://osf.io/6mwyk/

We start with 'script for figures and plots v5'

NB This assumes we already have the file summary_allNov14_2016.xlsx

# Script to produce information for Tables 1-3 and figs 2-3 (ERRNI data)
# Added to Github on 24 Jun 2020
###############################################################################
Script by Paul Thompson December2016
 Edited by Dorothy Bishop 21/12/16
 Added script to create Table 4 22/12/16  Included scripts for sequential analysis of proportions 22/12/16
```{r loadpackages}
#library(XLConnect) - original scripts used XLConnect to read xls files
#This is not working with current R version
library(openxlsx) #This added 2020 as alternative package for reading xlsx files
library(reshape2)
library(doBy)
library(ggplot2)
library(tidyr)
library(dplyr)
library(lme4)
library(car)
```

User will need to set paths to location of the raw files.  
NB the final raw data folder should ONLY contain included raw data files. The script will list all files in that folder and load them sequentially.  
The script also expects a folder ERRNI outputs where outputs will be stored.

```{r setpaths}
thisperson='dorothybishop' #change to pthompson if need be
mymac=1 #set to 0 if not using a mac

myfolder=paste("C:\\Users\\",thisperson,"\\Dropbox\\ERRNIanalysis2016\\ERRNI raw DATA\\final_041116",sep="")
mybackfile=paste("C:\\Users\\",thisperson,"\\Dropbox\\ERRNIanalysis2016\\ERRNI outputs\\ID_backgds.xlsx",sep="")
myoutputs=paste("C:\\Users\\",thisperson,"\\Dropbox\\ERRNIanalysis2016\\ERRNI outputs",sep="")
mysep="\\"
if (mymac==1){
  thisperson='dorothybishop'
  mysep="/"
  myfolder=paste("/Users/",thisperson,"/Dropbox/ERRNIanalysis2016/ERRNI raw DATA/final_041116",sep="")
  mybackfile=paste("/Users/",thisperson,"/Dropbox/ERRNIanalysis2016/ERRNI outputs/ID_backgds.xlsx",sep="")
  myoutputs=paste("/Users/",thisperson,"/Dropbox/ERRNIanalysis2016/ERRNI outputs",sep="")
}

location_final<-myfolder
myfinallist<-list.files(location_final) #NB the final raw data folder should ONLY contain included raw data files
Nsub=length(myfinallist)

setwd(myoutputs) #default is to write any newly created files to ERRNI outputs
```


```{r checkfiles}
#extract age bands
ageB <- substring(myfinallist,1,2)

#######################################

background_wb<-loadWorkbook(mybackfile)
#backgroundy=data.frame(readWorksheet(background_wb,sheet=1))
backgroundy=data.frame(readWorkbook(background_wb,sheet=1)) #version for openxlsx package

#######################################
#check files match
match(backgroundy$ID,substring(toupper(myfinallist),1,7))


#Before running code below select which story to run for "Fish or "beach"
id_ly<-data.frame(IDold=myfinallist, ID=substring(toupper(myfinallist),1,7))

id_er<-merge(id_ly,backgroundy,by="ID", all.x=T)

id_er<-id_er[is.na(id_er$story)==FALSE,]

```
# Table 1

```{r populate.table1}


# create a blank table in R to populate with values
tab1<-matrix(NA,14,6)
tab1<-as.data.frame(tab1)

#add table names (these are updated later)
names(tab1)<-c("Age_Band", "N", "Beach_fish","Male_Female","TROG","No._recall")

#populate column for age bands and N.
tab1$Age_Band<-names(table(id_er$ageband))
tab1$N<-table(id_er$ageband)

#Run a loop to fill values for beach|fish counts, Male|Female counts, and TROG mean(SD). 
for(i in 1:14){
  
  tab1$Beach_fish[i]<-paste(round(table(id_er$story[id_er$ageband==names(table(id_er$ageband))[i]])/sum(table(id_er$story[id_er$ageband==names(table(id_er$ageband))[i]]))*100),collapse = "|")
  #
  tab1$Male_Female[i]<-paste(round(table(id_er$sex[id_er$ageband==names(table(id_er$ageband))[i]])[2:1]/sum(table(id_er$sex[id_er$ageband==names(table(id_er$ageband))[i]])[2:1])*100),collapse = "|")
  
  tab1$TROG[i]<-paste(round(mean(id_er$TROG[id_er$ageband==names(table(id_er$ageband))[i]],na.rm=T),1)," (",round(sd(id_er$TROG[id_er$ageband==names(table(id_er$ageband))[i]],na.rm=T),1),")",sep = "")
}
```

NB for next chunk, I was getting some errors:
Error in match(sheetrId, file_rIds) : object 'sheetrId' not found.
I remembered that these sometimes seemed confused between xlsx and xls.
To open the raw file, need to change to xls and then resave as xlsx.

This affected several files, e.g.; 0751x01, 0635H01
I think these were ones that Pauline had processed: something incompatible about these.

```{r no_narrative}

#THis chunk then creates a list of the participants that did not produce a narrative
recaller<-matrix(NA,ncol=2,nrow=Nsub)
recaller<-as.data.frame(recaller)
names(recaller)<-c("ID","N_recall")
for(i in 1:Nsub){
  
  
  myID=paste(location_final,myfinallist[i],sep=mysep)
  
  mytranscript=loadWorkbook(myID)

  mytrans=data.frame(readWorkbook(mytranscript,sheet=1)) #changed from readWorksheet
  mytrans$N.word=as.numeric(mytrans$N.word)#needed because in some files this is non-numeric!!
  myagg=aggregate(mytrans$N.word~mytrans$TELL.RECALL, data=mytrans, FUN=function(x) c(mean=mean(x), count=length(x)))
  norecall=0
  aggrow=nrow(myagg)
  if (aggrow<2){norecall=1} #exclude cases with zero recall
  numbit=myagg[,2]#this is a matrix, not a number!
  if(sum(numbit[,2])<10) {norecall=1} #  exclude cases with fewer than total of 10 utterances
  if(mean(numbit[,1])<1.5){norecall=1} # exclude cases with MLU < 1.5 (only labelling)
  recaller[i,1] <-substring(myfinallist[i],1,7)
  recaller[i,2]<-norecall
}
```

```{r changecase}
#The ID columns are changed to upper case font so that they can be merged correctly.
id_er$ID<-as.vector(unlist(lapply(id_er$ID, function(v) {
  if (is.character(v)) return(toupper(v))
  else return(v)
})))

recaller$ID<-as.vector(unlist(lapply(recaller$ID, function(v) {
  if (is.character(v)) return(toupper(v))
  else return(v)
})))
```
#Merge data together, so that a summary column of percentage of "no narrative" participants by age can be added to Table 1.

```{r merge}
id_er2<-merge(id_er,recaller,by="ID")

for(i in 1:14){
  
  tab1$No._recall[i]<-sum(id_er2$N_recall[id_er2$ageband==names(table(id_er2$ageband))[i]],na.rm=T)
}


```

#Reorder table 1 so that younger age bands appear first.
```{r reordertab}
tab1<-tab1[c(9:14,1:8),]
names(tab1)<-c("Age_Band", "N", "%Beach_fish","%Male_Female","TROG ss","N no narrative")
#Write csv to file with data for table.
write.csv(tab1,"table_1_errni.csv",row.names=F)

```

# Pirate plots Fig 1
 NB loaded current version of yarrr in development via Github

```{r makepirate}
# Load data table

summary_dat_wb<-loadWorkbook("~/Dropbox/ERRNIanalysis2016/ERRNI outputs/summary_allNov14_2016.xlsx")
summary_dat=data.frame(readWorkbook(summary_dat_wb,sheet=1))

#Load R package to produce pirate plots.
#install.packages("yarrr")
library("yarrr")

summary_dat$ageband<-factor(summary_dat$ageband,levels=c( "4A", "4B", "4C", "4D", "5A", "5B","06", "07", "08", "09", "10", "12", "14", "17"))

allexclude=which(summary_dat$allexclude==1)
summary_dat=summary_dat[-allexclude,]
tiff(file = "Figure 1_col.tiff", width = 3200, height = 3200, units = "px", res = 500)
#par(xpd=T) #turned this off to avoid having distribution lines extending beyond axes

pcol<-c("skyblue","dodgerblue","dodgerblue3","dodgerblue4","green","green4","pink","yellow","red","purple","orange","deeppink4","darkmagenta","brown")

pirateplot(formula = avgmlu ~ ageband,
           theme=0,
           data = summary_dat,
           xlab = "Age Band",
           ylab = "MLUw",
           main = "",
           point.col=pcol,point.bg = pcol,
           inf.f.o = .7,
           avg.line.o = 1,
           bar.f.o = .01,
           bean.b.o = .1,
           point.o = .2,
           point.pch = 16,
           back.col = gray(.99))
# NB commands for this version of yarrr needed changing:
#hdi.o to inf.f.o, line.o to avg.line.o , and bar.o to bar.f.o
#dev.off()

```
