---
title: "ERRNI Notebook"
output: html_notebook
---

Notebook version of original analysis scripts for Frizelle et al.
The original scripts were uploaded to OSF around Dec 2016 https://osf.io/6mwyk/

NB from DVB, June 25 2020. This is pretty clunky!
It is particularly inefficient because different stages repeatedly read in all the raw data files.  
The sensible approach would be to read them in once, and then create a giant .csv file with all cases appended, after selecting the key columns (note that the columns in the original data files are consistently labelled, but inconsistent in terms of column numbering). This would create a classic longform file that could then be saved and reloaded for specific analyses.  

I have also appended a short chunk showing automation of MLU count.

# Original scripts
We start with 'script for figures and plots v5'

These are largely unmodified from originals, except I had problems with using XLConnect to read xlsx files with current version of R, so have substituted openxlsx.  

In general, though, I'd strongly recommend saving files as .csv, because it we've found that importing direct from xlsx may not work on all machines or with all R versions.  

NB The script assumes we already have the file summary_allNov14_2016.xlsx, which details background information - age group etc.

# Script to produce information for Tables 1-3 and figs 2-3 (ERRNI data)
Added to Github on 24 Jun 2020

Script by Paul Thompson December2016
Edited by Dorothy Bishop 21/12/16
Added script to create Table 4 22/12/16  Included scripts for sequential analysis of proportions 22/12/16
Modified by DVMB 24/06/20
```{r loadpackages}
#library(XLConnect) - original scripts used XLConnect to read xls files
#This is not working with current R version
library(openxlsx) #This added 2020 as alternative package for reading xlsx files
library(reshape2)
library(doBy)
library(ggplot2)
library(tidyr)
library(dplyr)
library(lme4)
library(car)
```

User will need to set paths to location of the raw files.  
NB the final raw data folder should ONLY contain included raw data files. The script will list all files in that folder and load them sequentially.  
The script also expects a folder ERRNI outputs where outputs will be stored.

```{r setpaths}
thisperson='dorothybishop' #change to pthompson if need be
mymac=1 #set to 0 if not using a mac

myfolder=paste("C:\\Users\\",thisperson,"\\Dropbox\\ERRNIanalysis2016\\ERRNI raw DATA\\final_041116",sep="")
mybackfile=paste("C:\\Users\\",thisperson,"\\Dropbox\\ERRNIanalysis2016\\ERRNI outputs\\ID_backgds.xlsx",sep="")
myoutputs=paste("C:\\Users\\",thisperson,"\\Dropbox\\ERRNIanalysis2016\\ERRNI outputs",sep="")
mysep="\\"
if (mymac==1){
  thisperson='dorothybishop'
  mysep="/"
  myfolder=paste("/Users/",thisperson,"/Dropbox/ERRNIanalysis2016/ERRNI raw DATA/final_041116",sep="")
  mybackfile=paste("/Users/",thisperson,"/Dropbox/ERRNIanalysis2016/ERRNI outputs/ID_backgds.xlsx",sep="")
  myoutputs=paste("/Users/",thisperson,"/Dropbox/ERRNIanalysis2016/ERRNI outputs",sep="")
}

location_final<-myfolder
myfinallist<-list.files(location_final) #NB the final raw data folder should ONLY contain included raw data files
Nsub=length(myfinallist)

setwd(myoutputs) #default is to write any newly created files to ERRNI outputs
```


```{r checkfiles}
#extract age bands
ageB <- substring(myfinallist,1,2)

#######################################

background_wb<-loadWorkbook(mybackfile)
#backgroundy=data.frame(readWorksheet(background_wb,sheet=1))
backgroundy=data.frame(readWorkbook(background_wb,sheet=1)) #version for openxlsx package

#######################################
#check files match
match(backgroundy$ID,substring(toupper(myfinallist),1,7))


#Before running code below select which story to run for "Fish or "beach"
id_ly<-data.frame(IDold=myfinallist, ID=substring(toupper(myfinallist),1,7))

id_er<-merge(id_ly,backgroundy,by="ID", all.x=T)

id_er<-id_er[is.na(id_er$story)==FALSE,]

```
# Table 1

```{r populate.table1}


# create a blank table in R to populate with values
tab1<-matrix(NA,14,6)
tab1<-as.data.frame(tab1)

#add table names (these are updated later)
names(tab1)<-c("Age_Band", "N", "Beach_fish","Male_Female","TROG","No._recall")

#populate column for age bands and N.
tab1$Age_Band<-names(table(id_er$ageband))
tab1$N<-table(id_er$ageband)

#Run a loop to fill values for beach|fish counts, Male|Female counts, and TROG mean(SD). 
for(i in 1:14){
  
  tab1$Beach_fish[i]<-paste(round(table(id_er$story[id_er$ageband==names(table(id_er$ageband))[i]])/sum(table(id_er$story[id_er$ageband==names(table(id_er$ageband))[i]]))*100),collapse = "|")
  #
  tab1$Male_Female[i]<-paste(round(table(id_er$sex[id_er$ageband==names(table(id_er$ageband))[i]])[2:1]/sum(table(id_er$sex[id_er$ageband==names(table(id_er$ageband))[i]])[2:1])*100),collapse = "|")
  
  tab1$TROG[i]<-paste(round(mean(id_er$TROG[id_er$ageband==names(table(id_er$ageband))[i]],na.rm=T),1)," (",round(sd(id_er$TROG[id_er$ageband==names(table(id_er$ageband))[i]],na.rm=T),1),")",sep = "")
}
```

NB for next chunk, I was getting some errors:
Error in match(sheetrId, file_rIds) : object 'sheetrId' not found.
I remembered that these sometimes seemed confused between xlsx and xls.
To open the raw file, need to change to xls and then resave as xlsx.

This affected several files, e.g.; 0751x01, 0635H01
I think these were ones that Pauline had processed: something incompatible about these.

```{r no_narrative}

#THis chunk then creates a list of the participants that did not produce a narrative
recaller<-matrix(NA,ncol=2,nrow=Nsub)
recaller<-as.data.frame(recaller)
names(recaller)<-c("ID","N_recall")
for(i in 1:Nsub){
  
  
  myID=paste(location_final,myfinallist[i],sep=mysep)
  
  mytranscript=loadWorkbook(myID)
  
  mytrans=data.frame(readWorkbook(mytranscript,sheet=1)) #changed from readWorksheet
  mytrans$N.word=as.numeric(mytrans$N.word)#needed because in some files this is non-numeric!!
  myagg=aggregate(mytrans$N.word~mytrans$TELL.RECALL, data=mytrans, FUN=function(x) c(mean=mean(x), count=length(x)))
  norecall=0
  aggrow=nrow(myagg)
  if (aggrow<2){norecall=1} #exclude cases with zero recall
  numbit=myagg[,2]#this is a matrix, not a number!
  if(sum(numbit[,2])<10) {norecall=1} #  exclude cases with fewer than total of 10 utterances
  if(mean(numbit[,1])<1.5){norecall=1} # exclude cases with MLU < 1.5 (only labelling)
  recaller[i,1] <-substring(myfinallist[i],1,7)
  recaller[i,2]<-norecall
}
```

```{r changecase}
#The ID columns are changed to upper case font so that they can be merged correctly.
id_er$ID<-as.vector(unlist(lapply(id_er$ID, function(v) {
  if (is.character(v)) return(toupper(v))
  else return(v)
})))

recaller$ID<-as.vector(unlist(lapply(recaller$ID, function(v) {
  if (is.character(v)) return(toupper(v))
  else return(v)
})))
```
#Merge data together, so that a summary column of percentage of "no narrative" participants by age can be added to Table 1.

```{r merge}
id_er2<-merge(id_er,recaller,by="ID")

for(i in 1:14){
  
  tab1$No._recall[i]<-sum(id_er2$N_recall[id_er2$ageband==names(table(id_er2$ageband))[i]],na.rm=T)
}


```

#Reorder table 1 so that younger age bands appear first.
```{r reordertab}
tab1<-tab1[c(9:14,1:8),]
names(tab1)<-c("Age_Band", "N", "%Beach_fish","%Male_Female","TROG ss","N no narrative")
#Write csv to file with data for table.
write.csv(tab1,"table_1_errni.csv",row.names=F)

```

# Pirate plots Fig 1
NB loaded current version of yarrr in development via Github

```{r makepirate}
# Load data table

summary_dat_wb<-loadWorkbook("~/Dropbox/ERRNIanalysis2016/ERRNI outputs/summary_allNov14_2016.xlsx")
summary_dat=data.frame(readWorkbook(summary_dat_wb,sheet=1))

#Load R package to produce pirate plots.
#install.packages("yarrr")
library("yarrr")

summary_dat$ageband<-factor(summary_dat$ageband,levels=c( "4A", "4B", "4C", "4D", "5A", "5B","06", "07", "08", "09", "10", "12", "14", "17"))

allexclude=which(summary_dat$allexclude==1)
summary_dat=summary_dat[-allexclude,]
tiff(file = "Figure 1_col.tiff", width = 3200, height = 3200, units = "px", res = 500)
#par(xpd=T) #turned this off to avoid having distribution lines extending beyond axes

pcol<-c("skyblue","dodgerblue","dodgerblue3","dodgerblue4","green","green4","pink","yellow","red","purple","orange","deeppink4","darkmagenta","brown")

pirateplot(formula = avgmlu ~ ageband,
           theme=0,
           data = summary_dat,
           xlab = "Age Band",
           ylab = "MLUw",
           main = "",
           point.col=pcol,point.bg = pcol,
           inf.f.o = .7,
           avg.line.o = 1,
           bar.f.o = .01,
           bean.b.o = .1,
           point.o = .2,
           point.pch = 16,
           back.col = gray(.99))
# NB commands for this version of yarrr needed changing:
#hdi.o to inf.f.o, line.o to avg.line.o , and bar.o to bar.f.o
#dev.off()

```

The next section incorporates a different script that was used to for table 3 and 4.

# Script for table 3 and 4 (ERRNI paper)

```{r clauseinfo}
#Gather all information about clauses from each excel file.
#This makes a long table with all IDs and clause codes
clauser<-matrix(NA,ncol=2)
clauser<-as.data.frame(clauser)
names(clauser)<-c("adding","sheet")
for(i in 1:length(myfinallist)){
  
  myID=paste(location_final,'/',myfinallist[i],sep="")
  if(file.exists(myID)){
    mytranscript=loadWorkbook(myID)
    mytrans=data.frame(readWorkbook(mytranscript,sheet=1))
    
    mytrans<-mytrans[!mytrans$clause %in% c("a+","cr+","cn+","n+","r+","cf+","t+"),]
    
    mytrans$clause<-recode(mytrans$clause,"'a-'='a';'cr-'='cr';'cf-'='cf';'u'=NA;'d-'='d';'n-'='n'")
    
    adding <- levels(factor(mytrans$clause))
    sheet <- rep(myfinallist[i],length(adding))
    clause1<-cbind(adding, as.character(sheet))
    colnames(clause1)<-c("adding","sheet")
    clauser<-rbind(clauser,clause1)
  }
  
}
#clauser1<-as.data.frame(clauser)
table(clauser$adding)

#Determine names of all available clauses in data
clauses<-names(table(clauser$adding))

#reorder names of clauses
clauses<-clauses[c(11,6,7,8,9,1,2,3,4,10,5)]

#set number of clauses
N.clause<-dim(table(clauser$adding))
```


Next chunk creates a count of all clauses from each participan by Tell and Recall.
Chunk also counts an instance of each clause used by each participant.
```{r clausecount}
clausecount<-matrix(NA,ncol=N.clause+2,nrow=length(myfinallist))
clausecount_T<-clausecount_R<-clausecount_all<-as.data.frame(clausecount)
clauseinst_T<-clauseinst_R<-clauseinst_all<-as.data.frame(clausecount)
for(i in 1:length(myfinallist)){
  
  myID=paste(location_final,'/',myfinallist[i],sep="")
  if(file.exists(myID)){
    mytranscript=loadWorkbook(myID)
    mytrans=data.frame(readWorkbook(mytranscript,sheet=1))
    
    #mytrans$clause <- factor(mytrans$clause)
    clausecount_T[i,1]<-clausecount_R[i,1]<-clausecount_all[i,1]<-substring(myfinallist[i],1,7)
    clausecount_T[i,2]<-clausecount_R[i,2]<-clausecount_all[i,2]<-substring(myfinallist[i],1,2)
    #
    clauseinst_T[i,1]<-clauseinst_R[i,1]<-clauseinst_all[i,1]<-substring(myfinallist[i],1,7)
    clauseinst_T[i,2]<-clauseinst_R[i,2]<-clauseinst_all[i,2]<-substring(myfinallist[i],1,2)
    
    mytrans<-mytrans[!mytrans$clause %in% c("a+","cr+","cn+","n+","r+","cf+","t+"),]
    
    mytrans$clause<-recode(mytrans$clause,"'a-'='a';'cr-'='cr';'cf-'='cf';'u'=NA;'d-'='d';'n-'='n'")
    
    for(j in 1:N.clause)
    {
      clauseinst_T[i,j+2]<-ifelse(match(clauses[j],names(table(mytrans$clause[mytrans$TELL.RECALL=="T"])),nomatch=0)>0,1,0)
      clauseinst_R[i,j+2]<-ifelse(match(clauses[j],names(table(mytrans$clause[mytrans$TELL.RECALL=="R"])),nomatch=0)>0,1,0)
      clauseinst_all[i,j+2]<-ifelse(match(clauses[j],names(table(mytrans$clause)),nomatch=0)>0,1,0)
      #
      clausecount_T[i,j+2]<-ifelse(match(clauses[j],names(table(mytrans$clause[mytrans$TELL.RECALL=="T"])),nomatch=0)>0,table(mytrans$clause[mytrans$TELL.RECALL=="T"])[names(table(mytrans$clause[mytrans$TELL.RECALL=="T"]))==clauses[j]][[1]],0)
      clausecount_R[i,j+2]<-ifelse(match(clauses[j],names(table(mytrans$clause[mytrans$TELL.RECALL=="R"])),nomatch=0)>0,table(mytrans$clause[mytrans$TELL.RECALL=="R"])[names(table(mytrans$clause[mytrans$TELL.RECALL=="R"]))==clauses[j]][[1]],0)
      clausecount_all[i,j+2]<-ifelse(match(clauses[j],names(table(mytrans$clause)),nomatch=0)>0,table(mytrans$clause)[names(table(mytrans$clause))==clauses[j]][[1]],0)
      
    }
    
  }
  
}


#Rename column headings for new summary tables of clauses counts and clause instances.
colnames(clausecount_T)<-c("ID","AGE_band",paste(clauses,"_T"))
colnames(clausecount_R)<-c("ID","AGE_band",paste(clauses,"_R"))
colnames(clausecount_all)<-c("ID","AGE_band",clauses)
#################################################################
colnames(clauseinst_T)<-c("ID","AGE_band",paste(clauses,"_T"))
colnames(clauseinst_R)<-c("ID","AGE_band",paste(clauses,"_R"))
colnames(clauseinst_all)<-c("ID","AGE_band",clauses)

#Function recodes the ID column so that all ID characters are upper case.

clausecount_T$AGE_band<-clauseinst_T$AGE_band<-as.vector(unlist(lapply(clausecount_T$AGE_band, function(v) {
  if (is.character(v)) return(toupper(v))
  else return(v)
})))
########
clausecount_R$AGE_band<-clauseinst_R$AGE_band<-as.vector(unlist(lapply(clausecount_R$AGE_band, function(v) {
  if (is.character(v)) return(toupper(v))
  else return(v)
})))
########
clausecount_all$AGE_band<-clauseinst_all$AGE_band<-as.vector(unlist(lapply(clausecount_all$AGE_band, function(v) {
  if (is.character(v)) return(toupper(v))
  else return(v)
})))
########
```
Next chunk calculates the means of clauses for each participant on their TELL responses.

```{r meanclauses}
clausemean_T<-matrix(NA,ncol=N.clause+1,nrow=length(unique(clausecount_T$AGE_band)))
clausemean_T<-as.data.frame(clausemean_T)
names(clausemean_T)<-c("AGE_band",clauses)

for(i in 1:length(unique(clausecount_T$AGE_band)))
{
  sub<-clausecount_T[clausecount_T$AGE_band==unique(clausecount_T$AGE_band)[i],]
  clausemean_T[i,2:(N.clause+1)]<-colMeans(sub[,-c(1,2)])
  clausemean_T[i,1]<-unique(clausecount_T$AGE_band)[i]
}

#calculates the means of clauses for each participant on their RECALL responses.

clausemean_R<-matrix(NA,ncol=N.clause+1,nrow=length(unique(clausecount_R$AGE_band)))
clausemean_R<-as.data.frame(clausemean_R)
names(clausemean_R)<-c("AGE_band",clauses)

for(i in 1:length(unique(clausecount_R$AGE_band)))
{
  sub<-clausecount_R[clausecount_R$AGE_band==unique(clausecount_R$AGE_band)[i],]
  clausemean_R[i,2:(N.clause+1)]<-colMeans(sub[,-c(1,2)])
  clausemean_R[i,1]<-unique(clausecount_R$AGE_band)[i]
}


#calculates the means of clauses for each participant on their combined TELL and RECALL responses.

clausemean_all<-matrix(NA,ncol=N.clause+1,nrow=length(unique(clausecount_all$AGE_band)))
clausemean_all<-as.data.frame(clausemean_all)
names(clausemean_all)<-c("AGE_band",clauses)

for(i in 1:length(unique(clausecount_all$AGE_band)))
{
  sub<-clausecount_all[clausecount_all$AGE_band==unique(clausecount_all$AGE_band)[i],]
  clausemean_all[i,2:(N.clause+1)]<-colMeans(sub[,-c(1,2)])
  clausemean_all[i,1]<-unique(clausecount_all$AGE_band)[i]
}

#Reorders the summary tables so that age bands are in descending order.

clausemean_R<-clausemean_R[c(9:14,1:8),]
clausemean_T<-clausemean_T[c(9:14,1:8),]
clausemean_all<-clausemean_all[c(9:14,1:8),]

################################################################################################
```

Calculates percentage of participants producing at least one utterance (TELL)
```{r percents}
clauseprop_T<-matrix(NA,ncol=N.clause+1,nrow=length(unique(clauseinst_T$AGE_band)))
clauseprop_T<-as.data.frame(clauseprop_T)
names(clauseprop_T)<-c("AGE_band",clauses)

for(i in 1:length(unique(clauseinst_T$AGE_band)))
{
  sub<-clauseinst_T[clauseinst_T$AGE_band==unique(clauseinst_T$AGE_band)[i],]
  clauseprop_T[i,2:(N.clause+1)]<-round((colSums(sub[,-c(1,2)])/dim(sub)[1])*100,2)
  clauseprop_T[i,1]<-unique(clauseinst_T$AGE_band)[i]
}

#Calculates percentage of participants producing at least one utterance (RECALL)

clauseprop_R<-matrix(NA,ncol=N.clause+1,nrow=length(unique(clauseinst_R$AGE_band)))
clauseprop_R<-as.data.frame(clauseprop_R)
names(clauseprop_R)<-c("AGE_band",clauses)

for(i in 1:length(unique(clauseinst_R$AGE_band)))
{
  sub<-clauseinst_R[clauseinst_R$AGE_band==unique(clauseinst_R$AGE_band)[i],]
  clauseprop_R[i,2:(N.clause+1)]<-round((colSums(sub[,-c(1,2)])/dim(sub)[1])*100,2)
  clauseprop_R[i,1]<-unique(clauseinst_R$AGE_band)[i]
}


#Calculates percentage of participant producing at least one utterance (ALL)

clauseprop_all<-matrix(NA,ncol=N.clause+1,nrow=length(unique(clauseinst_all$AGE_band)))
clauseprop_all<-as.data.frame(clauseprop_all)
names(clauseprop_all)<-c("AGE_band",clauses)

for(i in 1:length(unique(clauseinst_all$AGE_band)))
{
  sub<-clauseinst_all[clauseinst_all$AGE_band==unique(clauseinst_all$AGE_band)[i],]
  clauseprop_all[i,2:(N.clause+1)]<-round((colSums(sub[,-c(1,2)])/dim(sub)[1])*100,2)
  clauseprop_all[i,1]<-unique(clauseinst_all$AGE_band)[i]
}

#Reorders the summary tables so that age bands are in descending order.

clauseprop_R<-clauseprop_R[c(9:14,1:8),]
clauseprop_T<-clauseprop_T[c(9:14,1:8),]
clauseprop_all<-clauseprop_all[c(9:14,1:8),]
```

################################################################################################

 Write data tables to .csv file

```{r writefiles}
#Clause means

write.csv(clausemean_T,"clausemean_tell_table.csv",row.names = FALSE)
write.csv(clausemean_R,"clausemean_recall_table.csv",row.names = FALSE)
write.csv(clausemean_all,"clausemean_table.csv",row.names = FALSE)

################################################################################

# Clause proportions

write.csv(clauseprop_T,"clauseinstance_tell_table.csv",row.names = FALSE)
write.csv(clauseprop_R,"clauseinstance_recall_table.csv",row.names = FALSE)
write.csv(clauseprop_all,"clauseinstance_table.csv",row.names = FALSE)

################################################################################
```

# Word count/MLU

I've added here a bit of script to demonstrate doing a word count in R to get MLU.


We'll just take a sample file in .csv format to show how it works.
This has already had word count done, but we create another column, wordsN2, to redo it.  

Warning: the word count simply counts blanks; it may give errors if you have double blanks, or if words are demarcated some other way (e.g. if alternative versions of a word are both indicated).  

```{r readfile}
mydir<-"~/Dropbox/ERRNIanalysis2016/ERRNI R scripts/MLU scripts"#demofile should be in this directory

    filename<-paste0(mydir,'/demofile.csv')
    mytrans<-read.csv(filename,stringsAsFactors=F)
    
    #create wordsN2 as check on Nwords
    mytrans$wordsN2 <- sapply(strsplit(mytrans$text, " "), length)
  
    #word count only for legitimate clauses; NA for others
    legitlist <- c('m','t','a,','c','n','r','d') 
    thisclauselist <- substring(mytrans$clause,1,1) #this takes first character of clause
    #i.e. we include here subordinate clauses coded - and + in the word count
    w <- which(thisclauselist %in% legitlist)
    mytrans$wordsN2[-w]<-NA
    
    #For clause count, we only want main clauses
    w2 <- which(mytrans$clause %in% legitlist )
    
    MLU <- sum(mytrans$wordsN2,na.rm=T)/length(w2)
    
```

   